COLLISION RATES OF PLANETESIMALS NEAR MEAN-MOTION RESONANCES

### 2020-06-17 main_aaron.tex
Updated intro and made comments farther down in the manuscript.  Still working on it.

 \ACBc{[Feel free to omit this, particularly because it has a self citation, but I think it is important to highlight that your results could be relevant beyond debris systems]  While the focus here is on the evolution of planetesimal disks in general, with the most direct application being debris systems, we note that the results may be relevant for interpreting structures observed in young, gas-rich disks. For example, assuming giant planet formation requires first the presence of planetesimals, then systems such as HL Tau that have purported embedded giant planets, could have localized regions of dust enhancement due to the corresponding planetesimal collisions  \citep{2017ApJ...850..103B}.  The main difference between such cases and debris disks is that the dust in the former situation couples to gas. }

@@

\noindent Here, $\alpha = a_{g} / a$ where $a_{g}$ and $a$ are the semi-major axes of the giant planet and the planetesimal, 
respectively. $e_{g}$ and $\varpi_{g}$ are the eccentricity and longitude of pericenter of the giant and $b^{j}_{s} (\alpha)$ is a 
Laplace coefficient given by \citep{2000ssd..book.....M} as

\begin{equation}\label{eq:lap}
	b_{s}^{j}(\alpha) = \frac{1}{2 \pi} \int_{0}^{2 \pi} \frac{cos \, j \theta \, d \theta}{\left( 1 - 2 \alpha \, cos \theta + \alpha^2 \right)^{s}}.
\end{equation}

@@

In regions where there are commensurabilities between frequencies, secular theory breaks down and bodies are subject to strong 
perturbations. For the purposes of this study, we will ignore secular resonances, which generally occur on rather large timescales 
and will focus on mean motion resonances (MMRs). A MMR occurs  when the orbital period ratio between two bodies is sufficiently 
close to

@@

\noindent where  $p$ and $q$ are integers $>$ 0 and the unprimed and primed quantities correspond to the perturber and the body 
being perturbed, respectively. If the perturber is much more massive than the other body, the condition for MMR is set  by the semi-major 
axes of the two bodies by

@@

\noindent where $\lambda = \varpi + M$ is the mean longitude of a body. For bodies in resonance, the critical angle will librate around an 
equlibrium value, while this angle will circulate outside of resonance. For small eccentricities, this behavior is analogous to the motion of a 
pendulum. Variations in the critical angle are coupled to changes in the mean motion and semimajor axis \citet{2000ssd..book.....M}. An important 
point to note, which will become important later, is that the variation frequency of this angle approaches zero near the edge of a resonance. The 
width of a resonance can be defined by determining the largest variation in semimajor axis that permits librational, rather than circulatory motion of 
the $\phi$. For second and higher order resonances, the maximum libration width is given by

@@

The net result of this is that MMRs will produce spikes in the semimajor axis - eccentricity distribution of a planetesimal disk. In order to 
preserve the Jacobi energy, equation \ref{eq:tiss} predicts that these spikes will tend to curve away from the perturbing body. This lowers the 
surface density of the disk near the locations of resonances, while producing a pileup of bodies on eccentric orbits adjacent to the MMR. The 
enhanced surface density will increase the collision rate. At the same time, the large eccentricites will either lower it due to the decreased 
effectiveness of gravitational focusing, or raise it due to the more vigorous encounter rate, depending on the size of the encounter velocity relative to 
the escape velocity of the planetesimals. In addition, the MMRs cause the orbits of planetesimals to precess on relatively short 
timescales, which effectively randomizes their longitudes of perihelion. In a secularly forced disk, there will be an interface between aligned 
and randomly oriented orbits at the edge of the resonance. Due to the complicated dynamics that are at play in this region, we turn to a direct 
N-body treatment to better understand how the collision rate varies in the vicinity of a MMR.

@@

In total, five simulations are run. The first is a 'nominal' case, in which the perturbing planet's mass and eccentricity are set to that of Jupiter's. 
In the other four cases, the mass (eccentricity) is increased or decreased by a factor of two, while the eccentricity (mass) is held at the nominal 
value. In all cases, the perturbing giant is placed on a 5.2 AU orbit around a 1 $M_{\odot}$ star. The planetesimal disk extends from 2.2 to 3.8 
AU, which covers the two most prominent mean motion resonances with the giant planet, the 2:1 at 3.27 AU and the 3:1 at 2.5 AU. The 
planetesimal disk follows a minimum-mass solar nebula surface density profile \citep{1981PThPS..70...35H}

@@

\noindent and a forced component is added to $h$ according to equation \ref{eq:eforced} (where we have set $\varpi_{g}$ = 0).

@@

An additional timescale is set by the dynamical time of the planetesimals ($\sim 1/\sqrt{G \rho}$), which is about 45 minutes. To 
resolve the base time step and the dynamical timescale of planetesimals simultaneously, we use a two-tiered time stepping scheme. 
To start, all bodies are placed on the orbital time step. A first pass of collision detection is then run in 
which the radii of all bodies are inflated by a factor of 2.5. Any bodies with imminent collisions predicted using the inflated radii are 
placed on a time step that is a factor of 16 smaller than the orbital time step. The purpose of two-tiered scheme is to properly resolve 
the gravitational interactions between any bodies that undergo a close encounter. This prevents the coarser base time step from 
reducing the effectiveness of gravitational focusing, while minimizing the additional computational expense.

\begin{figure*}
\begin{center}
    \includegraphics[width=\textwidth]{figures/xy.png}
    \caption{\ACBc{fix aspect ratino so a circle is a circle.}  The positions of the remaining planetesimals at the end of the e1m2 (left), e2m2 (center) and e3m2 (right) simulations. The red dot 
    indicates the position of the giant planet and the dashed line points in the direction of the planet's longitude of perihelion. Non-axisymmetric gaps 
    are apparent near the locations of MMRs. At higher eccentricities, more resonances become visible. At the 2:1 MMR, gap features at $\theta$ = 0 
    and $\theta = \pi$ follow the giant planet in its orbit.\label{fig:xy}}
\end{center}
\end{figure*}

@@

We begin by examining simulations e1m2, e2m2 and e3m2. The positions of the planetesimals in the x-y plane after 5,000 years of integration are 
shown in figure \ref{fig:xy}. In all cases, the coordinate system is rotated so that the giant planet lies at $\theta = 0$. The longitude of perihelion of 
the planet is shown by the dashed line. Resonances with the perturbing planet are visible as nonaxisymmetric gaps. Upon close inspection of a 
series of simulation snapshots, these gaps appear to follow the planet in its orbit, rather than aligning themselves with the longitude of perihelion. A 
similar substructure, which is most noticeable near the 2:1 MMR, reveals itself in \citet{2000Icar..143...45R} (see bottom panel of figure 3) and 
\citet{2016ApJ...818..159T} (figure 3). It is worth noting that \citet{2000Icar..143...45R} started with a completely cold planetesimal disk and a Jupiter 
mass planet on a circular, coplanar orbit. The presence of these features seems robust to the choice of initial conditions.

@@

\begin{figure}
\begin{center}
    \includegraphics[width=0.5\textwidth]{figures/coll_hist_a.png}
    \caption{In semimajor axis space, prominent features appear near the 3:1 and 2:1 MMRs. Near the 3:1, the collision
    rate decreases from its unperturbed value, while an enhancement apepars near the 2:1.\label{fig:coll_hist_a}}
\end{center}
\end{figure}

>>>>

 \ACBc{[UPDATED to take into account further discussion later in the article. Much of the discussion above is on debris discs, so I am trying to reframe the discussion for what follows later in the manuscript. Feel free to omit this, particularly because it has a self citation, but I think it is important because the intro is a bit off point otherwise.]  The focus here is on the collisional evolution of planetesimal disks in general, and the results are not intended to be confined to debris disks.   During planet formation, which occurs in the presence of planetesimals and gas, the coupling between gas and collisionally-produced dust could lead to observable signatures \citep{2017ApJ...850..103B}, as we will explore.  }

@@

\noindent Here, $\alpha = a_{g} / a$ where $a_{g}$ and $a$ are the semi-major axes of the giant planet and the planetesimal, 
respectively. $e_{g}$ and $\varpi_{g}$ are the eccentricity and longitude of pericenter of the giant and $b^{j}_{s} (\alpha)$ is a 
Laplace coefficient given by \citep{2000ssd..book.....M} as \ACBc{[fixed $cos$ to $\cos$.]}

\begin{equation}\label{eq:lap}
	b_{s}^{j}(\alpha) = \frac{1}{2 \pi} \int_{0}^{2 \pi} \frac{\cos \, j \theta \, d \theta}{\left( 1 - 2 \alpha \, \cos \theta + \alpha^2 \right)^{s}}.
\end{equation}

@@

In regions where there are commensurabilities between frequencies, \ACBc{Laplace-Langrange} secular theory breaks down and bodies are subject to strong 
perturbations. For the purposes of this study, we will ignore secular resonances, which generally occur on rather large timescales 
and will focus on mean motion resonances (MMRs). A MMR occurs  when the orbital period ratio between two bodies is sufficiently 
close to

@@

\noindent where  $p$ and $q$ are integers $>$ 0 and the unprimed and primed quantities correspond to the perturber and the body 
being perturbed, respectively. \ACBc{[Might as well state the order and degree of the resonance to fully define the equation.]} If the perturber is much more massive than the other body, the condition for MMR is set  by the semi-major 
axes of the two bodies by

@@

\noindent where $\lambda = \varpi + M$ is the mean longitude of a body. For bodies in resonance, the critical angle will librate around an 
equlibrium value, while this angle will circulate outside of resonance. For small eccentricities, this behavior is analogous to the motion of a 
pendulum. Variations in the critical angle are coupled to changes in the mean motion and semimajor axis \citep{2000ssd..book.....M}\ACBc{[fixed citation]}. An important 
point to note, which \sout{will become important}\ACBc{we will come back to} later, is that the variation frequency of this angle approaches zero near the edge of a resonance. The 
width of a resonance can be defined by determining the largest variation in semimajor axis that permits librational, rather than circulatory motion of 
the $\phi$. For second and higher order resonances, the maximum libration width is given by

@@

The net result of this is that MMRs will produce spikes in the semimajor \ACBc{axis-eccentricity} distribution of a planetesimal disk. In order to 
preserve the Jacobi energy, equation \ref{eq:tiss} predicts that these spikes will tend to curve away from the perturbing body. This lowers the 
surface density of the disk near the locations of resonances, while producing a pileup of bodies on eccentric orbits adjacent to the MMR. The 
enhanced surface density will increase the collision rate. At the same time, the large eccentricites will either lower it due to the decreased 
effectiveness of gravitational focusing, or raise it due to the more vigorous encounter rate, depending on the size of the encounter velocity relative to 
the escape velocity of the planetesimals. In addition, the MMRs cause the orbits of planetesimals to precess on relatively short 
timescales, which effectively randomizes their longitudes of perihelion. In a secularly forced disk, there will be an interface between aligned 
and randomly oriented orbits at the edge of the resonance. Due to the complicated dynamics that are at play in this region, we turn to a direct 
N-body treatment to better understand how the collision rate varies in the vicinity of a MMR.

@@

In total, five simulations are run\ACBc{[Reference table!]}. The first is a `nominal' \ACBc{[fixed quotes]} case, in which the perturbing planet's mass and eccentricity are set to that of Jupiter's. 
In the other four cases, the mass (eccentricity) is increased or decreased by a factor of two, while the eccentricity (mass) is held at the nominal 
value. In all cases, the perturbing giant is placed on a 5.2 AU orbit around a 1 $M_{\odot}$ star\ACBc{, although the overall dynamics presented here are scalable to different scales [true, right?]}. The planetesimal disk extends from 2.2 to 3.8 
AU, which covers the two most prominent mean motion resonances with the giant planet, the 2:1 at 3.27 AU and the 3:1 at 2.5 AU. The 
planetesimal disk follows a minimum-mass solar nebula surface density profile \citep{1981PThPS..70...35H}

@@

\noindent and a forced component is added to $h$ according to equation \ref{eq:eforced} (where we have set $\varpi_{g}$ = 0). \ACBc{[question and minor.  Doesn't the convention place the cosine with k or q (for inclination).  I tend to recall this because the mnemonic device is that the hard sound goes with the cosine. I do not have my standard references on me to verify!  Feel free to ignore me.]}

@@

An additional timescale is set by the dynamical time of the planetesimals ($\sim 1/\sqrt{G \rho}$), which is about 45 minutes. To o i
resolve the base time step and the dynamical timescale of planetesimals simultaneously, we use a two-tiered time stepping scheme. 
To start, all bodies are placed on the orbital time step. A first pass of collision detection is then run in 
which the radii of all bodies are inflated by a factor of 2.5. Any bodies with imminent collisions predicted using the inflated radii are 
placed on a time step that is a factor of 16 smaller than the orbital time step. The purpose of \ACBc{the} two-tiered scheme is to properly resolve 
the gravitational interactions between any bodies that undergo a close encounter. This prevents the coarser base time step from 
reducing the effectiveness of gravitational focusing, while minimizing the additional computational expense.

\begin{figure*}
\begin{center}
    \includegraphics[width=\textwidth]{figures/xy.png}
    \caption{\ACBc{[Is this aspect ratio right? I.e., is a circle a circle?]}  The positions of the remaining planetesimals at the end of the e1m2 (left), e2m2 (center) and e3m2 (right) simulations. The red dot 
    indicates the position of the giant planet and the dashed line points in the direction of the planet's longitude of perihelion. Non-axisymmetric gaps 
    are apparent near the locations of MMRs. At higher eccentricities, more resonances become visible. At the 2:1 MMR, gap features at $\theta$ = 0 
    and $\theta = \pi$ follow the giant planet in its orbit.\label{fig:xy}}
\end{center}
\end{figure*}

@@

We begin by examining simulations e1m2, e2m2 and e3m2. The positions of the planetesimals in the x-y plane after 5,000 years of integration are 
shown in figure \ref{fig:xy}. In all cases, the coordinate system is rotated so that the giant planet lies at $\theta = 0$. The longitude of perihelion of 
the planet is shown by the dashed line. Resonances with the perturbing planet are visible as nonaxisymmetric gaps. Upon close inspection of a 
series of simulation snapshots, these gaps appear to follow the planet in its orbit, rather than aligning themselves with the longitude of perihelion. A 
similar substructure, which is most noticeable near the 2:1 MMR, reveals itself in \citet{2000Icar..143...45R} (see bottom panel of figure 3) and 
\citet{2016ApJ...818..159T} (figure 3). \ACBc{This structure was also seen in the \citet{2017ApJ...850..103B} simulations, although it was not reported at the time (see Appendix BLAH).}  It is worth noting that \citet{2000Icar..143...45R} started with a completely cold planetesimal disk and a Jupiter 
mass planet on a circular, coplanar orbit. \ACBc{ \citet{2017ApJ...850..103B} used a similar setup.  Thus, t}he presence of these features seems robust to the choice of initial conditions.

@@

\begin{figure}
\begin{center}
    \includegraphics[width=0.5\textwidth]{figures/coll_hist_a.png}
    \caption{In semimajor axis space, prominent features appear near the 3:1 and 2:1 MMRs. Near the 3:1, the collision
    rate decreases from its unperturbed value, while an enhancement apepars near the 2:1. \ACBc{[Suggest state bin width and mention density estimator -- basically, many people will just look at the figures, so some of this information is appropriate to repeat. ]} .\label{fig:coll_hist_a}} 
\end{center}
\end{figure}
